---
title: "Biostats-202C-hw1"
author: "Jiahao Tian"
date: "2022-10-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyr)
library(ggplot2)
library()
```

**Question 1**


**a)**

$$ y \sim pois(\lambda), ~~~\lambda \sim gamma(a, b) $$
$$ f(y|\lambda) = {(\lambda^y e^{-\lambda})\over y!} ~~~ L(y|\lambda) = {{\lambda^{\sum(yi)} e^{-ny}} \over \Pi y!} $$
$$ p(\lambda) = {b^a \over \Gamma(a)} \lambda^{a-1} e^{-\lambda{b}} $$
$$ p(\lambda | y) \propto \lambda^{\sum(yi)+a-1} e^{-\lambda(b+n)}, \lambda >0 $$
$$ posterior~ follows~ gamma~ distribution~ \sim Gamma(\sum(yi)+a, b+n) $$

**b)**

$$\begin{aligned}
f(y) & = \int_{0}^{\infty}{b^a \over \Gamma(a)}{\lambda^{a-1}}{e^{-\lambda b}}{{\lambda^ye^{-\lambda}}\over y!}d\lambda \\
& = {b^a \over {\Gamma(a)y!}}\int_{0}^{\infty}{\lambda^{a-1}}{{e^{-\lambda b}}}{\lambda^{y}}{e^{-\lambda}}d\lambda\\
& = {b^a \over {\Gamma(a)y!}}{\Gamma(y+a) \over (b+a)^{y+a}} \int_{0}^{\infty}{(b+a)^{y+a} \over \Gamma(y+a)}{\lambda^{y+a-1}}{e^{-\lambda(b+1)}}d\lambda \\
& = {b^a \over {\Gamma(a)y!}}{\Gamma(y+a) \over (b+1)^{y+a}}
\end{aligned}$$


**c)**

From part b,

$$ \begin{aligned}
f(y|a,b) & = {b^a \over \Gamma(a)y!}{{\Gamma(y+a)} \over (b+1)^{y+a}} \\
& = {(y+a-1)! \over (a-1)!y!}{b^a \over (b+1)^{y+a}}\\
& = {y+a-1 \choose y}b^a(b+1)^{-y-a}
\end{aligned}$$

$$ Which ~~ f(y|a,b) \sim NB(a,b)$$ 

**d)**

```{r}
y <- 1
data <- dpois(y, 1)
prior <- dgamma(y, 1, 1)
posterior <- dgamma(y, 2, 2)

plot(y, prior, type = "l", col = "red")
lines(y, posterior, type = "l", col = "green")
lines(y, data , type = "l", col = "blue")
```


**Question 2**


**a)**


$$L(y|\lambda) = {{\lambda^{\sum(yi)} e^{-ny}} \over \Pi y!} ~~~ p(\lambda) = {b^a \over \Gamma(a)} \lambda^{a-1} e^{-\lambda{b}} $$
$$ p(\lambda | y) \propto \lambda^{\sum(yi)+a-1} e^{-\lambda(b+n)}, \lambda >0 $$

$$ posterior~ follows~ gamma~ distribution~ \sim Gamma(\sum(yi)+a, b+n) $$

**Question 3**


**a)**

$$ \begin{aligned}
mean; ~~ E(y|\tau) & = \int_{0}^{\infty}y{({2 \over {\pi\tau}})^{1/2}}e^{-(y)^2 \over 2\tau}dy \\
& = ({2 \over \pi\tau})^{1/2}[-\tau e^{-{{y}^2}\over2\tau}|_{0}^{\infty}] \\
& = ({2\tau \over \pi})^{1/2}
\end{aligned}$$
$$ variance; ~~ \tau(1-{2 \over \pi})~~~~~~mode; ~~ 0~~~~~median; ~~ \tau\sqrt{2}{erf}^{-1}(1/2)$$
**b)**


$$ indicator,~~~~I = \{^{1~~~~0~<~yi}_{0~~~~otherwise}$$
$$\begin{aligned} Lf(y|\tau) & = {({2 \over {\pi\tau}})^{n/2}}e^{-{(\sum yi)^2 \over 2\tau}}I \\
logLf(y|\tau) & = {n \over 2}log{2 \over {\pi\tau}}-{{\sum yi}^2 \over {2\tau}}I \\
\triangledown logLf(y|\tau) & = -{n\over {2\tau^3}}+{{\sum yi^2}\over{2\tau^2}}~,set~equal~to~0\\
we~get~,\hat{\tau} & = {n \over {{\sum yi}}^2}\\
T & = {n \over {{\sum yi}}^2}~,is ~ the~ sufficient~ statistics ~for ~\tau 
\end{aligned}$$



**c)**


$$ indicator,~~~~I = \{^{1~~~~0~<~yi}_{0~~~~otherwise}$$


$$Lf(y|\tau) = {({2 \over {\pi\tau}})^{n/2}}e^{-{(\sum yi)^2 \over 2\tau}}I $$
$$f(\tau) = {{({b \over 2})^{a \over 2}} \over {\Gamma({a \over 2})}}\tau^{-{a \over 2}-1}e^{-{b \over {2\tau}}}  $$
$$ posterior;~~ f(\tau|y) \propto ({2 \over {\pi \tau^{a+2}}})^{n \over 2} e^{-{{({\sum yi}^2}+b}) \over 2\tau},~yi>0$$
$$posterior ~follows~ halfnormal ~distribution$$



**Question 4**


**a)**

When data is non-negative, also, collected at the individual level (binary response) or group level (proportion) might use
half-normal distribution. When data shows binary response and non-negative values will lead to use half-normal distribution.


**b)**

Example: Half-normal can be use to model distance at which wildlife observe on a boat detect an animal when surveying, such 
as whales in the ocean.

**c)**

The competitor distribution I might use is "Poisson distribution".



**Question 5**


**a)**

$$ f(y|\sigma) = {({2 \over {\pi\sigma}})^{1/2}}e^{-{(y)^2 \over 2\sigma}},~y>0  $$

**b)**

Yes, because half-normal distribution's variance here is unknown and inverse gamma can use as the conjugate prior of the 
variance.


**c)**

Yes, I thinks that it can be inverse gamma.



**Question 6**


**a)**

$$ f(y|\theta) \sim beta(\theta,1) $$
$$ so~the~normalizing~constant~is;~ {{\Gamma(\theta +1)} \over \Gamma(\theta) \Gamma(1)}$$


**b)**



$$ Lf(y|\theta) = \bigg({{\Gamma(\theta +1)} \over \Gamma(\theta) \Gamma(1)}\bigg)^n\Pi_{i=1}^n {yi}^{\theta - 1}I_{0<yi<1}$$

$$\begin{aligned}
By~Factorization \\
h(y) & = \Pi_{i=1}^n {yi}^{- 1} \\
k(g(y)|\theta) & = \Pi_{i=1}^n {yi}^{\theta}I_{0<yi<1} \\
T & = max \{yi\}, is ~ the~ sufficient~ statistics ~for ~\theta
\end{aligned}$$






















